---
title: "HW2"
output:
  html_document:
    df_print: paged
---


# Previous steps

1.  Create GitHub repo for this homework.
2.  Clone the repo locally.
3.  Prepare your code for the homework on that repo.
4.  Download the file "hollywood.graphml"
5.  You can open the file locally using.

```{r}
set.seed(1111)
rm(list = ls())
library(igraph)

dataGitLink='https://github.com/DACSS-CSSmeths-summer/datafiles/raw/main/hollywood.graphml'
actors=read_graph(dataGitLink,format='graphml')
```

As you see, you have:

```{r}
summary(actors)
```

From above:

- The graph "actors" is an undirected network of actors.
```{r}
V(actors)
```
- The actor node has an attribute _wonOscar_, if they won an oscar in acting.
```{r}
V(actors)$wonOscar # 0 means NO, 1 yes.
```
Maybe better if we relabel
```{r}
V(actors)$wonOscar=ifelse(V(actors)$wonOscar==1,'Yes','No')
```

- There is an edge between two actors are connected if they have ever acted in the same film. 
```{r}
E(actors)
```


- The network is **weighted** (edge attribute), that is, the link/edge informs how many times they have been together in a film.
```{r}
E(actors)$weight # most actors worked together once.
```

You can see the network here:

```{r, message=FALSE,warning=FALSE}
library(ggraph)
ggraph(graph = actors) + 
  geom_edge_link() + 
  geom_node_point()
```

You add the weights as labels and edge width:

```{r, message=FALSE,warning=FALSE}
ggraph(graph = actors) + 
  geom_edge_link(aes(label = weight,
                     edge_width =  weight),
                 color='grey70' ) + 
  geom_node_point() 
```

You modify the theme:

```{r, message=FALSE,warning=FALSE}
ggraph(graph = actors) + 
  geom_edge_link(aes(label = weight ),
                 color='grey70' ) + 
  geom_node_point() + theme_graph() +
  geom_node_text(
    aes(label = name))
```

This could be a final version:
```{r, message=FALSE,warning=FALSE, fig.width=10, fig.height=6}
finalGraph=ggraph(graph = actors) + 
  geom_edge_link(aes(edge_width = as.ordered(weight) ),
                 color='grey70' ) + 
  geom_node_point() + theme_graph() +
  geom_node_text(aes(label = name,color = as.factor(wonOscar)),
                 size=5)

## here the final version
finalGraph
```

Maybe better?

```{r, warning=FALSE}

## here the final version
finalGraph + labs(edge_width="Movies in commmon",
                  color="Won Oscar as actor") 

```


# TO DO

1.  Read from YOUR OWN GitHub (cloud): **dataGitLink** above should point to your repo (not mine)
2.  Compute node central measures: eigen_centrality, closeness, betweenness.
3.  Prepare a correlation plot with previous measures, and **comment** the result.
4.  Redo the plot `finalGraph`, just showing the Hub nodes.
5.  Answer: Is it possible communities are present? Support your answer with computations.
6.  IF the presence of communities is possible, compute the communities using `Girvan-Newman` (aka edge_betweenness), and `Louvain` algorithms, or other algorithms. Plot only the best result.

